Re-analysing the data from @moffatt_inner_2020: A textbook illustration of the absence of evidence fallacy

### 2021-04-11 manuscript.Rmd
minor updates to manuscript

The activity of silently talking to oneself or "inner speech" is a foundational ability, allowing oneself to remember, plan, self-motivate or self-regulate [for reviews, see @alderson-day_inner_2015; @perrone-bertolotti_what_2014; @loevenbruck_cognitive_2018]. However, whereas the use of inner speech is associated with many adaptive functions in everyday life, inner speech dysfunctions can be identified in multiple psychological disorders. For instance, rumination, broadly defined as unconstructive repetitive thinking about past events and current mood states [@Martin], is involved in the onset and maintenance of serious mental disorders such as depression, anxiety, eating disorders or substance abuse [for a review, see @Nolen-Hoeksema2008].

Given the predominantly verbal nature of rumination [e.g., @ehring_repetitive_2008; @goldwin_concreteness_2012; @goldwin_concreteness_2013; @mclaughlin_effects_2007], we previously proposed to consider rumination as a form of inner speech and to study it using the methods that have been used to study other forms of inner speech, namely, by using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_dissociating_2020; @nalborczyk_articulatory_2020]. We first showed that induced rumination was accompanied by increased facial (both over a forehead and a perioral site) muscular activity in comparison to a rest period [@nalborczyk_orofacial_2017]. However, because rumination was only compared to a rest period, it remained uncertain whether this perioral activity was specifically related to (inner) speech processes. Therefore, we ran a follow-up study comparing verbal to non-verbal rumination, which suggested that the facial EMG correlates we had previously identified were not specifically related to the verbal content of the ruminative thoughts [@nalborczyk_dissociating_2020]. We discussed these findings in length and proposed several theoretical interpretations that can account for these results in the discussion section of @nalborczyk_dissociating_2020 and more extensively in @nalborczyk_understanding_2019. Although these points were not addressed by @moffatt_inner_2020, their experimental design nevertheless had the potential to inform our understanding of the involvement of the speech motor system in different varieties of inner speech as well as to clarify the relation between the peripheral correlates of inner speech and (self-reported) subjective experience.

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to *severe* tests [e.g., @mayo_severe_2006; @mayo_statistical_2018]. In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives *severe* tests. More precisely, some claim (e.g., $\theta = 0$) is said to be *severely tested* if it had great chances of being corroborated/falsified, had the claim been true/false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the *relative evidence* for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we simulated 10.000 datasets (for $N = 26$) under the assumption of either no effect (i.e., the null hypothesis of $d = 0$), an effect size of $d = 0.36$ [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of $d = 0.72$ [i.e., the effect size reported in @nalborczyk_orofacial_2017].

@@

As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis (although the "true" effect size is $d = 0$). When the "true" effect size is of $d = 0.36$, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of $d = 0.72$, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates. To assess the extent to which the BF computed for the OOI by @moffatt_inner_2020 (i.e., $\text{BF}_{10} = 0.278$) is "surprising" given or "compatible" with some effect size, we can compute the proportion of simulated BFs that are equal or more extreme than the one reported by the authors for each hypothesis, which is approximately equal to `r round(1 - mean(results$BF10_null > 0.278), 2)`, `r round(1 - mean(results$BF10 > 0.278), 2)`, and `r round(1 - mean(results$BF10_double > 0.278), 2)`, for the hypothesis of $d = 0$, $d = 0.36$, and $d = 0.72$, respectively. In other words, observing a $\text{BF}_{10} = 0.278$ is more compatible with the hypothesis of $d = 0$ (i.e., the null hypothesis) than with the hypothesis of $d = 0.36$ or $d = 0.72$ (at this sample size).

@@

Another possible interpretation is that automatic forms of inner speech may rely more heavily on higher-level (e.g., memory-based) cognitive processes whereas less automatic (i.e., more intentional or deliberate) forms of inner speech may rely more on simulation mechanisms via the use of internal models of the speech motor system [@nalborczyk_understanding_2019]. In other words, the production of automatic versus non-automatic inner speech would be underpinned by different processes that would involve the speech motor system to a different extent. This distinction is similar to the distinction between the two routes of predictions-by-association and prediction-by-simulation in speech perception and comprehension [@pickering_integrated_2013]. The prediction-by-association mechanism would rely more on perceptual sensory experiences and domain-general cognitive abilities whereas the prediction-by-simulation mechanism would rely more on the simulation of the motor action leading to the speech auditory percept. In the former case, no peripheral muscular activity is expected, whereas in the latter case, the speech motor system would be involved in simulating or emulating the corresponding overt action [cf. also the motor simulation vs. direct simulation distinction in @tian_mental_2012]. Whether the physiological correlates of automatic versus non-automatic (deliberate) forms of inner speech differ because of inhibitory constraints or because they rely on different processes (e.g., prediction-by-association or prediction-by-simulation) remains an open empirical question. We previously discussed these issues in more length and suggested ways forward from an experimental perspective in the discussion of @nalborczyk_understanding_2019.

@@

# Supplementary materials {#supp}

Reproducible code and figures are available at https://osf.io/ba3gk/.

@@

>>>>

The activity of silently talking to oneself or "inner speech" is a foundational ability, allowing oneself to remember, plan, self-motivate or self-regulate [for reviews, see @alderson-day_inner_2015; @perrone-bertolotti_what_2014; @loevenbruck_cognitive_2018]. However, whereas the use of inner speech is associated with many adaptive functions in everyday life, inner speech dysfunctions can be identified in multiple psychological disorders. For instance, rumination, broadly defined as unconstructive repetitive thinking about past events and current mood states [@Martin], is involved in the onset and maintenance of serious mental disorders such as depression, anxiety, eating disorders, or substance abuse [for a review, see @Nolen-Hoeksema2008].

Given the predominantly verbal nature of rumination [e.g., @ehring_repetitive_2008; @goldwin_concreteness_2012; @goldwin_concreteness_2013; @mclaughlin_effects_2007], we previously proposed to consider rumination as a form of inner speech and to study it using the methods that have historically been used to study other forms of inner speech, namely, by using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_dissociating_2020; @nalborczyk_articulatory_2020]. We first showed that induced rumination was accompanied by increased facial (both over a forehead and a perioral site) muscular activity in comparison to a rest period [@nalborczyk_orofacial_2017]. However, because rumination was only compared to a rest period, it remained uncertain whether this perioral activity was specifically related to (inner) speech processes. Therefore, we ran a follow-up study comparing verbal to non-verbal rumination, which suggested that the facial EMG correlates we had previously identified were not specifically related to the verbal content of the ruminative thoughts [@nalborczyk_dissociating_2020]. We discussed these findings in length and proposed several theoretical interpretations that can account for these results in the discussion section of @nalborczyk_dissociating_2020 and more extensively in @nalborczyk_understanding_2019. Although these points were not addressed by @moffatt_inner_2020, their experimental design nevertheless had the potential to inform our understanding of the involvement of the speech motor system in different varieties of inner speech as well as to clarify the relation between the peripheral correlates of inner speech and (self-reported) subjective experience.

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to *severe* tests [e.g., @mayo_statistical_2018; @mayo_severe_2006]. In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives *severe* tests. More precisely, some claim (e.g., $\theta = 0$) is said to be *severely tested* if it had great chances of being corroborated/falsified, had the claim been true/false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the *relative evidence* for the null, computing a BF (i.e., comparing two models) does not solve the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we simulated 10.000 datasets (for $N = 26$) under the assumption of either no effect (i.e., the null hypothesis of $d = 0$), an effect size of $d = 0.36$ [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of $d = 0.72$ [i.e., the effect size reported in @nalborczyk_orofacial_2017].

@@

As shown in Figure \@ref(fig:bf-dance), the distribution of log-BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, `r round(prop_d0 * 100, 2)`% of the computed log-BFs are above 0 and hence support the alternative hypothesis (although the "true" effect size is $d = 0$). When the "true" effect size is of $d = 0.36$, `r round(100 - prop_d0.36 * 100, 2)`% of the log-BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of $d = 0.72$, `r round(100 - prop_d0.72 * 100, 2)`% of the log-BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates. To assess the extent to which the BF computed for the OOI by @moffatt_inner_2020 (i.e., $\text{BF}_{10} = 0.278$) is "surprising" given or "compatible" with some effect size, we can compute the proportion of simulated BFs that are equal or more extreme than the one reported by the authors for each hypothesis, which is approximately equal to `r round(1 - mean(results$BF10_null > 0.278), 2)`, `r round(1 - mean(results$BF10 > 0.278), 2)`, and `r round(1 - mean(results$BF10_double > 0.278), 2)`, for the hypothesis of $d = 0$, $d = 0.36$, and $d = 0.72$, respectively. In other words, observing a $\text{BF}_{10} = 0.278$ is more compatible with the hypothesis of $d = 0$ (i.e., the null hypothesis) than with the hypothesis of $d = 0.36$ or $d = 0.72$, for this sample size.

@@

Another possible interpretation is that automatic forms of inner speech may rely more heavily on higher-level (e.g., memory-based) cognitive processes whereas less automatic (i.e., more intentional or deliberate) forms of inner speech may rely more on simulation mechanisms via the use of internal models of the speech motor system [@nalborczyk_understanding_2019]. In other words, the production of automatic versus non-automatic inner speech would be underpinned by different processes that would involve the speech motor system to a different extent. This distinction is similar to the distinction between the two routes of prediction-by-association and prediction-by-simulation in speech perception and comprehension [@pickering_integrated_2013]. The prediction-by-association mechanism would rely more on perceptual sensory experiences and domain-general cognitive abilities whereas the prediction-by-simulation mechanism would rely more on the simulation of the motor action leading to the speech auditory percept. In the former case, no peripheral muscular activity is expected, whereas in the latter case, the speech motor system would be involved in simulating or emulating the corresponding overt action [cf. also the distinction between *motor simulation* and *direct simulation* in @tian_mental_2012]. Whether the physiological correlates of automatic versus non-automatic (deliberate) forms of inner speech differ because of inhibitory constraints or because they rely on different processes (e.g., prediction-by-association or prediction-by-simulation) remains an open empirical question. We previously discussed these issues in more length and suggested ways forward from an experimental perspective in the discussion of @nalborczyk_understanding_2019.

@@

@@

# Funding information

This project did not benefit from any specific funding.

# Data accessibility {#supp}

Reproducible code for the manuscript, analyses, and figures is available at [https://osf.io/ba3gk/](https://osf.io/ba3gk/).

# Authors contribution

LN, MPB, CB, RG, EK, and HL contributed to conception and design of the study, LN collected the data, LN, MPB, CB, EK, and HL contributed to data analysis and interpretation, LN, MPB, CB, RG, ES, EK, and HL drafted and/or revised the article and all authors approved the submitted version for publication.

# Competing Interests

The author has no competing interests to declare.
